Towards caring 15-minute neighbourhoods

### 2024-06-16 paper/Manuscript.qmd
Incorporated Lea and Lucia's comments

`r word_comment("We should focus this discussion on: defining mobility of care and connecting it to the 15-minute neighbourhoods, the importance of walk-accessibility, and socio-demographic/economic characteristics associated with this overlap (esp. single parent households, areas with a high proportion of children, low-income prevalence).", " ")`

Caring is one of the most fundamental activities in societies, which has allowed humanity to survive and evolve. It involves activities targeted to meet physical, psychological, and emotional needs of others [@ilo_care_2018]. Yet, it is one of the worst devalued, underpaid, and unequal activities worldwide. Traditionally, caring activities have remained on women’s shoulders. Even though the society has evolved, and nowadays women attain higher educational levels and tend to work out of home more than ever before, the bulk of care remains (at any stage of her lifecourse) heavily on her shoulders. The Second Job or the Second Shift remains in all industrial societies, mostly women’s work (Hayden, 1982 (?Citation)) [@hochschild_second_2012].

This unequal share of caring responsibilities leads to gendered differences in among others, career development, differences in contract types and decision-making positions, unbalanced share of women in certain professions, gender pay gap, time poverty, as recognised by different international organisations (European Institute for Gender Equality, 2016 (?Citation)) [@ilo_care_2018]. In fact, already the first feminists in the United States observed that domestic labor was the most relevant basic cause of women’s inequality [@gordon_womans_1976] (Hayden, 1982 (?Citation)).

According to the [@ilo_care_2018], women and girls perform more than three-quarters of the total amount of unpaid care work and two-thirds of care workers are women. The caring economy is part of the women-dominated industries, it is characterised by the evaluative discrimination, which is only one of the dimensions of the gender pay gap [@mejiadorantes_discussing_2019]. Especially in the case of domestic workers, care workers are most of the times women and migrants’ workers, with poor working conditions, and many times associated with informal jobs which imply lack of labour rights, no social and health protection, susceptible of exploitation, one of the poorest working conditions across the care economy as the ILO observes [@ilo_care_2018]. Care workers becomes disrespected, leading to a devaluation of their activity, and consequently remains as one of the less well-paid occupations a person can undertake.

If we turn our attention to the urban planning perspective, Caring is a major human activity that may take form in all urban settings, in all transport modes, all times a day and year, and yet it is not described as neither origin nor destination, and until recently, it was inexistent.

Researchers from various fields have highlighted that western modern industrial societies have imposed a physical separation of the private sphere from the public one, splitting production and reproduction, public and private spaces or simply work and home, concentrating on the economic outcome of being mobile [@cresswell_gendered_2008; @lawWomenTransportNew1999; @levy2013travel; @little_gender_1994; @tronto_toward_1990]. The man with legal rights, paid labour public sphere and woman in the private sphere, dependent, with family duties, personal relations and unpaid labour [@tronto_toward_1990]. In this line, @cresswell_gendered_2008 described how masculinity was related to mobile and active, whereas femininity to static. As an example of this dichotomy, transport planning historically framed non-work/school trips as 'non-mandatory' activities, downgrading these activities in comparison to work.

@@

From this research motivation, Sánchez de Madariaga coined the term ‘mobility of care’ in 2013 to refer to all travel required to sustain the needs of a household, such as grocery shopping, escorting children, travelling to health appointments, and running errands [@sanchezdemadariagaMobilityCareIntroducing2013]. While decades of research had examined these unique household-serving trips, her seminal work was the first to consider all these trips as one category. In doing so, she showed how mobility of care is a significant proportion of daily travel. Indeed, in her original work, and in the studies that have followed, it appears that approximately 30% of adults’ trips are for care purposes [@sanchezdemadariagaMobilityCareIntroducing2013; @sanchezdemadariagaMeasuringMobilitiesCare2019; @ravensbergen2023exploratory]. Sánchez de Madariaga not only shows how important these mobility of care trips are, but also highlights the ways in which “mobility of care is systematically under-represented in any analysis of urban transport” (p. 37). Transport surveys and tools do not directly capture mobility of care, which re-enforces the idea that these trips are not a significant part of daily mobility.

Cities have simplified individuals as a universal and neutral entity. As a result, public spaces are differently experienced depending on our multiple identities. Indeed, asymmetric in the distribution of time of citizens is deeply anchored with the public and private spaces organization within cities. In this light, we focus on the potential access to destinations of care through an empirical case study of the mid-sized city of Hamilton, Canada. We acknowledge that caring is composed by actors and activities but aim to spotlight the role that the urban environment plays on our daily lives, on easing our participation and on exercising our rights to the city as individuals with multiple identities. Our work hopes to contribute to a fruitful discussion that may trigger urban and transport policy interventions that increase gender-equity in caring activities among all actors.

<!-- We experience our urban environment depending on our personal characteristics, such as socioeconomic level, education, profession, household responsibilities, physical or mental disabilities, and very importantly, by gender. Furthermore, our living environment is also influenced by historical, economic, social, cultural, political, religion and environmental factors. All these issues have an influence on the way we live and experience the city, on how we participate in the community and on our social relations in the public and private space, on how we invest in our socio-economic development, and finally on our well-being.  -->

<!-- Cities have simplified individuals as a universal and neutral entity. As a result, public spaces are differently experienced depending on our multiple identities. Indeed, the distribution of time of citizens is deeply anchored with the public and private spaces organization within cities. In fact, the asymmetries in the distribution of time of individuals are rooted on city spaces, distances to reach them, transport alternatives and different perceived fears, emotions and constraints, along with the roles typically associated with men and women, together with the gendered division of labour [@villagran_urban_2020]. Furthermore, these authors emphasize that it is a global reality that women invest a greater amount of time to care and to unpaid work and that it is socioeconomically stratified. -->
<!-- Many studies worldwide have documented how gender influences mobility patterns. Men commute longer distances and present direct travel patterns, whereas women present chained trips to accomplish on their way many different activities (for example, dropping children off at schools or daycare, run errands or accompany family members to medical services). For these reasons it has been documented that women tend to have shorter but more complicated transport patterns. They also tend to rely more on public transport whereas men use more private transport [@cresswell_gendered_2008; @levy2013travel; @mejia2021mobility; @the_world_bank_handbook_2020]. Sánchez de Madariaga uses the concept “mobility of care” to recognize, measure, make visible, value and account for all the travel associated to caring and unpaid tasks for the household [@sanchezdemadariagaMobilityCareIntroducing2013]. -->

@@

@@

In Paris, the adoption of the 15-Minute city put into the spotlight this approach. In Paris, they targeted six social functions that had to be easily accessible at any part of the city: housing, work, health care, groceries, education and leisure [@ville_de_paris_paris_2022]. In 2021, Ottawa also adopted a 15-Minutes approach in the new City of Ottawa Official Plan [@ville_dottawa_quartier_2021].

More recently, @moreno_introducing_2021 considers four dimensions on how the 15-minute city framework: density (in terms of people per sq.km); diversity (including mixed-land use and diversity of people; the temporal and spatial proximity to essential services (residential areas, work, commercial areas, education centres, health facilities and other basic institutions) and digitalisation.

It can be observed that the feminist perspective of the cities of proximity is still underestimated. Only few examples exist that specifically target this issue [@gil_sola_choose_2022; @macintyre_her_2022].

Even if we are on the right direction to facilitate caring activities, develop human-centred urban areas, the chrono-urbanism needs to set the ground on aspects that specifically target caring activities and gender. Similarly, it is necessary to provide alternatives to measure progress on this important issue. On the other hand, an intersectional approach to understand how different personal identities interact in our daily day activities in the cities is necessary. There is still room for improvement.

@@

Given that almost one third of daily trips are for care purposes [@sanchezdemadariagaMobilityCareIntroducing2013], it is important to plan for mobility of care. There are also equity considerations to consider: as discussed, women, and especially low-income women, complete this travel more than men. Planning for mobility of care may also warrant different considerations because mobility of care trip characteristics differ from other trips, notably those of the commute to work. For instance, many of these trips are made through trip-chains, i.e., when a non-work-related stop is made during the commute. Women have been found to do so more than men, which is understood as being due to their need to balance unpaid-household serving care and paid employment (e.g., picking up groceries or children on the way home from work) [@ravensbergen2023exploratory; @scheinerWomenComplexDaily2015].

@@

It appears that the car is the preferred travel mode for mobility of care. In one study, car use was higher for mobility of care trips than the commute to work (82.3% vs. 74.7%) [@ravensbergen2023exploratory]. Further, the rate of car use is higher amongst trips that are part of a chain than those that are not (85.28% v. 78.09%) [@ravensbergen2023exploratory]. Some studies found that a car is viewed as the easier way to incorporate unpaid mobility of care into one’s daily travel as it provides convenience and mitigates concerns about traffic and stranger danger [@maciejewskaHaveChildrenThus2019, @carverParentalChauffeursWhat2013]. Others, however, have found that women are more likely to complete household-serving travel duties by alternative modes than men due to their lower car access [@RN24;@RN25]. It seems that car use may be preferred for mobility of care trips, but planning mobility of care for the car may leave those who do not have access to a car, many of them low-income women, with difficult mobility of care trips.

@@

RESEARCH OBJECTIVES:

1)	15 minute cities stem from conventional methods that ignore mobility of care perspectives. There is an applied way of doing this.
2)	Use of the innovative destination dataset
3)	Mid-size typical Canadian city case study. How frequently do we find these caring neighbourhoods (potential - accessibility)? Clustering methodology to determine this.
4)	Who enjoys them and who does not? Clustering methodology to determine these profiles.
  - i.e. with LICO > X + Prevalence of children >Y + walking accessibility score > U + diversity score > W = “caring neighbourhood”

@@

- Next: we describe Hamilton's land-use context; communities which are urban, rural, suburban. Also describe the residential parcels summary statistics, ~140,000 points, aggregated at the level of Canadian Census DA. See parcel, population and population per parcel plots in @fig-Fig1:
```{r plot-figure1, out.width=600, out.height=600}
#| label: fig-Fig1
#| fig-cap: "The number of residential parcels per DA in 2020 (top), the population (middle) retrieved from the 2021 Canadian Census, and the rate of population per parcel per DA (bottom). All scales in quartiles. Basemap shapefiles are sourced from the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/figures/pop_descriptive_plots_1.png"))
```
Some of the socio-economic and demographic variables that characterise accessibility and the diversity of accessibility to care destinations are pictured in @fig-Fig2:

@@

Here we will describe the estimated 15-minute walk travel times using {r5r}, from all residential parcel points to all care destinations assuming default walking speed (3.6km/hr) and OSM road network for Hamilton. Provide summary statistics.

@@

```{r creating-table-content}
dat <- data.frame(matrix(c("Child-centric",
                           "[@opendatahamiltonEducationalInstitutions2022; @governmentofontarioOntarioDataCatalogue2023; @cityofhamiltonChildCareRegistry2023; @opendatahamiltonParksDatabase2022; @opendatahamiltonRecreationCommunityCentres2022]",
                           "Schools, daycares, and community centres, recreation centres, and parks: 1,190 locations are included. After manual review, all locations that typically do not serve children were removed including: Post-Secondary, Adult-Learning Centres, Group Homes, and Foster Care Centres. Further, through examination some Section 23 institutions defined as _“centres for children who cannot attend school to meet the needs of care or treatment, and rehabilitation”_ [@governmentofontarioministryofeducationFundingEducationCommunity2023], were kept due to their innate connection to care.",
                           "Elder-centric",
                           "[@opendatahamiltonRecreationCommunityCentres2022; @ontarioministryofhealthgeohubMinistryHealthService2023]",
                           "Senior centres, long-term care homes, and retirement homes: 75 destinations are identified.",
                           "Grocery-centric",
                           "[@axledataConsumerData2023]",
                           "Grocery stores, namely a place a household could buy groceries ranging from convenience stores to large retail stores: 381 destinations are identified. Data is filtered by Company Name, Suite Number, Address, City, Province, Phone Number and Postal Code. The type was then identified e.g., grocers specialty foods, grocers retail, grocer health food, grocer wholesale, grocer curbside, grocer delicatessen wholesale, grocer convenience. Data was crossreferenced to ensure all included locations were operational and legitimate grocery stores.",
                           "Health-centric",
                           "[@ontarioministryofhealthgeohubMinistryHealthService2023; @hamiltonniagarahaldimandbranthealthlineWalkInMedicalClinics2023]",
                           "Hospitals, pharmacies, clinics, and dentist offices: 421 destinations are identified. Hospitals and pharmacies were retrieved while clinics and dentistry clincs were manually scraped from a healthcare services database and checked via Google Maps to remove non-operational locations and confirm dentistry-orientation. ",
                           "Errand-centric",
                           "Hamilton libraries [@opendatahamiltonLibraries2022], post office locations [@axledataConsumerData2023;@canadapostPostOfficeLocator2023], and datasets of all national bank chains [@bankofmontrealBankMontrealBranch2023; @thehongkongandshanghaibankingcorporationlimitedhsbcHSBCBranchABM2023; @nationalbankNationalBankBranches2023; @royalbankofcanadaRBCBranchABM2023; @scotiabankScotiabankABMBranch2023; @thetorontodominionbankTDBankBranch2023].",
                                                      "Libraries, post offices, and banks: 158 destinations are identified. Post offices are retrieved from a mix of databases, and duplicates are removed. Banks are also derived from Data Axle and then cross-referenced to ensure data quality with a Bank Locator website for all national banking firms."
                           ), ncol = 3, byrow = T))
```

@@

## Methods
Cumulative opportunity method; 15-minute normative catchment for walk mode is used to represent the 15-minute neighbourhood concept. This has been done in previous literature such as @vale2023accessibility, @graellsCityCitiesMeasuring2021 and @gaglioneUrbanAccessibility15minute2022.

In this work, the cumulative opportunity score is calculated for each parcel $i$ and aggregated by care category $c$ and takes following mathetical form:
$$
S^c_i=\sum_{j}W_j^{c}\cdot f(c_{ij})
$$
\noindent Where:

@@

*Entropy* measure has been widely used to characterise the diversity of land-use mix in early work by @frankLinkingObjectivelyMeasured2005 based on the species diversity Shannon-Wiener index that expresses relative evenness throughout a sequence (this measure among other species diversity indices of the time are reviewed in @whittaker1972evolution). In the realm of urban studies and planning literature, it has since been widely used to characterise the diversity in land-use mix [@ewingTravelBuiltEnvironment2010a], especially in the context of active modes [@luUrbanDensityDiversity2017; @mavoaIdentifyingAppropriateLanduse2018] and suburban sprawl [@randallGISBasedLand2015]. Work that has used diversity indices alongside accessibility analysis are scarce, though examples of its use as parameters within accessibility scores in the case of employment opportunities [@chengMeasuringUrbanJob2013; @daiIncorporatingJobDiversity2018] and more recently to describe the diversity in transit facilities [@yinIncorporatingFacilityDiversity2024] are present. To the authors knowledge, the use of an entropy measure to reflect the diversity in care destination amenity types and in context within a 15-minute neighbourhood, has yet to be done.

@@

Diversity @fig-Fig5:
```{r plot-figure5, out.width=600, out.height=600}
#| label: fig-Fig5
#| fig-cap: "The diversity measure based on the proportion of care category spatial accessibility (figure 4). Scale is in tertiles. The fifth input into the superclusters. Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/figures/all_cats_entropy_copp_plot.png")) 
```

@@

```{r plot-figure7, out.width=600, out.height=600}
#| label: fig-Fig7
#| fig-cap: "The decision tree profiles of the SOM output."
knitr::include_graphics(glue::glue(here::here(), "/figures/decision_tree_profiles_sc.png"))
```

@@

These findings related to potential child-care access are notable: single parents provide more care and hence may be more likely to choose, if they are able, more accessible neighbourhoods than the rest of the population in the same household income grouping. In a Canadian and U.S. context, households with children may locate to catchments with 'better' schools, often more suburban with lower potential access scores overall.

@@

>>>>

Caring is one of the most fundamental activities in societies, which has allowed humanity to survive and evolve. It involves activities targeted to meet the physical, psychological, and emotional needs of others [@ilo_care_2018]. Yet, it is one of the most devalued, underpaid, and unequal activities worldwide. Traditionally, caring activities have remained on women’s shoulders. Though women today attain higher educational levels and tend to work out of home more than ever before, the bulk of care remains (at any life stage) heavily on women's shoulders. The 'Second Shift' remains mostly women’s work  in all industrial societies [@hayden_grand_2000; @hochschild_second_2012]. According to the @ilo_care_2018, women and girls perform more than three-quarters of the total amount of unpaid care. This gender gap varies geographically [@RN5]. In Canada, adult women complete 1.9 hours, on average, more of unpaid care than men every day [@RN4].

This unequal share of caring responsibilities leads to gendered differences in: career development, profession selection, contract type, pay gap, and time poverty as recognized by different international organisations [@EIGE2016; @ilo_care_2018]. In fact, even early feminists in the United States observed that domestic labor was the most relevant basic cause of women’s inequality [@gordon_womans_1976; @hayden_grand_2000].

Care also has significant planning considerations. Indeed, researchers from various fields have highlighted that western modern industrial societies have imposed a physical separation of the private sphere from the public one, splitting production and reproduction, public and private spaces or simply work and home, concentrating on the economic outcome of being mobile [@cresswell_gendered_2008; @lawWomenTransportNew1999; @levy2013travel; @little_gender_1994; @tronto_toward_1990]. The man with legal rights, paid labour public sphere and woman in the private sphere, dependent, with family duties, personal relations and unpaid labour [@tronto_toward_1990]. In this line, @cresswell_gendered_2008 described how masculinity was related to mobile and active, whereas femininity to static. As an example of this dichotomy, transport planning historically framed non-work/school trips as 'non-mandatory' activities, downgrading these activities in comparison to work. In more recent years, planners have begun to resist this spatial separation of the public and private sphere to promote for sustainable cities. Proximity planning has become more normalized, at popular example of which is the 15-minute city, in which residents can reach all necessary destinations within a 15-minute walk. The 15-minute city concept, however, has yet to directly focus on care.

While much care labour takes place in the private, and feminised sphere, completing care activities requires one to be mobile. In other words: caring for oneself and others requires trips to, amongst other places, grocery stores, day cares, elder care, and health care destinations.

@@

From this research motivation, Sánchez de Madariaga coined the term ‘mobility of care’ in 2013 to refer to this travel or all travel required to sustain the needs of a household, such as grocery shopping, escorting children, travelling to health appointments, and running errands [@sanchezdemadariagaMobilityCareIntroducing2013]. While decades of research had examined these unique household-serving trips, her seminal work was the first to consider all these trips as one category. In doing so, she showed how mobility of care is a significant proportion of daily travel. Indeed, in her original work, and in the studies that have followed, it appears that approximately 30% of adults’ trips are for care purposes [@sanchezdemadariagaMobilityCareIntroducing2013; @sanchezdemadariagaMeasuringMobilitiesCare2019; @ravensbergen2023exploratory]. Perhaps unsurprisingly given the gendered division of care work discussed above, women have consistently been found to complete more mobility of care trips than men. In one study, mobility of care comprised 32% of women’s daily trips compared to 28% of men’s. While this gap was significant, it was found to be far greater in lower income households where women complete 20% more care trips than men [@ravensbergen2023exploratory]. Sánchez de Madariaga not only shows how important these mobility of care trips are, but also highlights the ways in which “mobility of care is systematically under-represented in any analysis of urban transport” (p. 37). Transport surveys and tools do not directly capture mobility of care, which re-enforces the idea that these trips are not a significant part of daily mobility.

Cities have simplified individuals as a universal and neutral entity. As a result, public spaces are differently experienced depending on our multiple identities. Indeed, asymmetric in the distribution of time of citizens is deeply anchored with the public and private spaces organization within cities. In this light, we bring together the literatures on 15-minute cities and mobility of care by focusing on the potential access to destinations of care through an empirical case study of the mid-sized city of Hamilton, Canada. We acknowledge that caring is composed by actors and activities but aim to spotlight the role that the urban environment plays on our daily lives, on easing our participation and on exercising our rights to the city as individuals with multiple identities. Our work hopes to contribute to a fruitful discussion that may trigger urban and transport policy interventions that increase gender-equity in caring activities among all actors.

@@

`r word_comment("Good 15 minute city example, but what about feminist 15 minute cities? It's novel but movements exists such as https://x.com/brenttoderian/status/1800371515877384527?s=48&t=QI0WSvV_FDb82ZGeLdJ9lQ", " ")`

@@

In Paris, the adoption of the 15-Minute city put into the spotlight this approach. They targeted six social functions that had to be easily accessible at any part of the city: housing, work, health care, groceries, education and leisure [@ville_de_paris_paris_2022]. In 2021, Ottawa also adopted a 15-Minutes approach in the new City of Ottawa Official Plan [@ville_dottawa_quartier_2021]. More recently, @moreno_introducing_2021 considers four dimensions on how the 15-minute city framework: density (in terms of people per sq.km); diversity (including mixed-land use and diversity of people; the temporal and spatial proximity to essential services (residential areas, work, commercial areas, education centres, health facilities and other basic institutions) and digitalisation.  In this respect, the feminist perspective of the cities of proximity is still underestimated. Only few examples exist that specifically target this issue [@gil_sola_choose_2022; @macintyre_her_2022]. Given the impact of caring activities on everyday life, the chrono-urbanism should specifically target the analysis of caring activities.

@@

Given that almost one third of daily trips are for care purposes [@sanchezdemadariagaMobilityCareIntroducing2013], it is important to plan for mobility of care. There are also equity considerations to consider: as discussed, women, and especially low-income women, are particularly in charge of these duties and therefore, complete this travel more than men. Planning for mobility of care may also warrant different considerations because mobility of care trip characteristics differ from other trips, notably those of the commute to work. For instance, many of these trips are made through trip-chains, i.e., when a non-work-related stop is made during the commute. Women have been found to do so more than men, which is understood as being due to their need to balance unpaid-household serving care and paid employment (e.g., picking up groceries or children on the way home from work) [@ravensbergen2023exploratory; @scheinerWomenComplexDaily2015].

@@

It appears that the car is the preferred travel mode for mobility of care. In one study, car use was higher for mobility of care trips than the commute to work (82.3% vs. 74.7%) [@ravensbergen2023exploratory]. Further, the rate of car use is higher amongst trips that are part of a chain than those that are not (85.28% vs. 78.09%) [@ravensbergen2023exploratory]. Some studies found that a car is viewed as the easier way to incorporate unpaid mobility of care into one’s daily travel as it provides convenience and mitigates concerns about traffic and stranger danger [@maciejewskaHaveChildrenThus2019, @carverParentalChauffeursWhat2013]. Others, however, have found that women are more likely to complete household-serving travel duties by alternative modes than men due to their lower car access [@RN24;@RN25]. It seems that car use may be preferred for mobility of care trips, but planning mobility of care for the car may leave those who do not have access to a car, many of them low-income women, with difficult mobility of care trips.

@@

The objectives of the paper are as follows:

1)	City planning is moving to a people-centered approach focusing on accessibility to basic needs. However, chrono-urbanism stems from conventional methods that overlooks mobility of care perspectives. This paper presents a 15-minute cities methodological approach to assess mobility of care.
2)	We develop this approach using a mid-size typical Canadian city and a novel care destination dataset as case study
3)	Furthermore, the potential accessibility to and diversity of care destinations in the case study is analysed. We identify to what degree are neighbourhoods caring and/or complete. To do so, we make use of self organizing maps (SOM), an unsupervised machine learning data-dimensionality reduction technique to identify caring and complete neighbourhoods.
4)	We then apply the same SOM methodology to determine profiles of population groups to understand who enjoys these neighbourhoods and who does not.

@@

This study focuses on Hamilton, Ontario, a mid-sized city on the shore of Lake Ontario. Hamilton has a heterogenous land-use, with a populated and dense urban core, surrounded by suburban development, which is itself surrounded by rural communities made possible through a greenbelt. In this work, we analyse the residential parcel centroids, approximately ~140,000 locations. We aggregated the points at the level of Canadian Census Dissemination Area (DA) along with the population and population per parcel plots in @fig-Fig1.

```{r plot-figure1, out.width=600, out.height=600}
#| label: fig-Fig1
#| fig-cap: "The number of residential parcels per DA in 2020 (top), the population (middle) retrieved from the 2021 Canadian Census, and the rate of population per parcel per DA (bottom). All scales in quartiles. Basemap shapefiles are sourced from the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/figures/pop_descriptive_plots_1.png"))
```

Hamilton also exhibits spatial disparities in social and economic indicators @fig-Fig2. The densely populated inner city is characterised by lower average incomes, and a higher prevalence of households living under the LICO. The suburban parts of the city tend to have a greater proportion of children and a lower proportion of one parent households.

@@

Here we will describe the estimated 15-minute walk travel times using {r5r}, from all residential parcel points to all care destinations assuming default walking speed (3.6km/hr) and OSM road network for Hamilton. **Provide summary statistics.**

@@

```{r creating-table-content}
dat <- data.frame(matrix(c("Child-centric",
                           "[@opendatahamiltonEducationalInstitutions2022; @governmentofontarioOntarioDataCatalogue2023; @cityofhamiltonChildCareRegistry2023; @opendatahamiltonParksDatabase2022; @opendatahamiltonRecreationCommunityCentres2022]",
                           "Schools, daycares, and community centres, recreation centres, and parks: 1,190 locations are included. After manual review, all locations that typically do not serve children were removed including: Post-Secondary, Adult-Learning Centres, Group Homes, and Foster Care Centres. Further, through examination some Section 23 institutions defined as _“centres for children who cannot attend school to meet the needs of care or treatment, and rehabilitation”_ [@governmentofontarioministryofeducationFundingEducationCommunity2023], were kept due to their innate connection to care.",
                           "Elder-centric",
                           "[@opendatahamiltonRecreationCommunityCentres2022; @ontarioministryofhealthgeohubMinistryHealthService2023]",
                           "Senior centres, long-term care homes, and retirement homes: 75 destinations are identified.",
                           "Grocery-centric",
                           "[@axledataConsumerData2023]",
                           "Grocery stores, namely a place a household could buy groceries ranging from convenience stores to large retail stores: 381 destinations are identified. Data is filtered by Company Name, Suite Number, Address, City, Province, Phone Number and Postal Code. The type was then identified e.g., grocers specialty foods, grocers retail, grocer health food, grocer wholesale, grocer curbside, grocer delicatessen wholesale, grocer convenience. Data was cross-referenced to ensure all included locations were operational and legitimate grocery stores.",
                           "Health-centric",
                           "[@ontarioministryofhealthgeohubMinistryHealthService2023; @hamiltonniagarahaldimandbranthealthlineWalkInMedicalClinics2023]",
                           "Hospitals, pharmacies, clinics, and dentist offices: 421 destinations are identified. Hospitals and pharmacies were retrieved while clinics and dentistry clinics were manually scraped from a healthcare services database and checked via Google Maps to remove non-operational locations and confirm dentistry-orientation. ",
                           "Errand-centric",
                           "Hamilton libraries [@opendatahamiltonLibraries2022], post office locations [@axledataConsumerData2023;@canadapostPostOfficeLocator2023], and datasets of all national bank chains [@bankofmontrealBankMontrealBranch2023; @thehongkongandshanghaibankingcorporationlimitedhsbcHSBCBranchABM2023; @nationalbankNationalBankBranches2023; @royalbankofcanadaRBCBranchABM2023; @scotiabankScotiabankABMBranch2023; @thetorontodominionbankTDBankBranch2023].",
                                                      "Libraries, post offices, and banks: 158 destinations are identified. Post offices are retrieved from a mix of databases, and duplicates are removed. Banks are also derived from Data Axle and then cross-referenced to ensure data quality with a Bank Locator website for all national banking firms."
                           ), ncol = 3, byrow = T))
```

@@

## Methods
A 15-minute normative walk catchment is used to represent the 15-minute neighbourhood concept. This has been done in previous literature such as @vale2023accessibility, @graellsCityCitiesMeasuring2021 and @gaglioneUrbanAccessibility15minute2022. In this work, the cumulative opportunity score is calculated for each parcel $i$ and aggregated by care category $c$ and takes the following mathematical form:
$$
S^c_i=\sum_{j}O_j^{c}\cdot f(c_{ij})
$$
\noindent Where:

@@

*Entropy* measure has been widely used to characterise the diversity of land-use mix in early work by @frankLinkingObjectivelyMeasured2005 based on the species diversity Shannon-Wiener index that expresses relative evenness throughout a sequence (this measure among other species diversity indices of the time are reviewed in @whittaker1972evolution). It has also been used in social sciences to comprehend the occurrence of social events in a structured manner specially with regard to life trajectories [@ritschardSequenceAnalysisRelated2018].

In the realm of urban studies and planning literature the malleable concept of entropy has been used to characterise the diversity in land-use mix [@ewingTravelBuiltEnvironment2010a], especially in the context of active modes [@luUrbanDensityDiversity2017; @mavoaIdentifyingAppropriateLanduse2018] and suburban sprawl [@randallGISBasedLand2015]. As well, it has been used to understand mobility behaviour [@mcbride2020exploration; @montero2023applying; @montero2023role]. Work that has used diversity indices alongside accessibility analysis are scarce, though examples of its use as parameters within accessibility scores in the case of employment opportunities [@chengMeasuringUrbanJob2013; @daiIncorporatingJobDiversity2018] and more recently to describe the diversity in transit facilities [@yinIncorporatingFacilityDiversity2024] are present. To the authors knowledge, the use of an entropy measure to reflect the diversity in care destination amenity types and in context within a 15-minute neighbourhood, has yet to be done.

@@

Diversity @fig-Fig5:
```{r plot-figure5, out.width=600, out.height=600}
#| label: fig-Fig5
#| fig-cap: "The diversity measure (div) based on the proportion of care category spatial accessibility (figure 4). Scale is in tertiles. The fifth input into the superclusters. Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/figures/all_cats_entropy_copp_plot.png")) 
```

@@

```{r plot-figure7, out.width=600, out.height=300}
#| label: fig-Fig7
#| fig-cap: "The decision tree profiles of the SOM output."
knitr::include_graphics(glue::glue(here::here(), "/figures/decision_tree_profiles_sc.png"))
```

@@

These findings related to potential child-care access are notable: single parents provide more care and hence may be more likely to choose, if they are able, more accessible neighbourhoods than the rest of the population in the same household income grouping. In a Canadian and U.S. context, households with children may locate to catchments with 'better' schools, often more suburban with lower potential access scores overall. Furthermore, statistics show that in the last years, the number of single parents families in Canada has increased. In 2022, there were about 1.84 million single parents compared to about 1.56 million in 2010 and 1,3 million in 2001 (https://www.statista.com/statistics/443342/single-parent-families-in-canada/ & https://www.justice.gc.ca/eng/rp-pr/jr/jr12/p5a.html#ftn4). Furthermore, Ontario is the Canadian province where most of the lone parent families live (749,769 families, https://www.statista.com/statistics/446667/lone-parent-families-in-canada-by-province/). Most lone-families are headed by female-lone parent. (AS TO ADD IN HAMILTON SPECIFIC NUMBERS)

@@

#Conclusions and Discussion

There seems to be a spatial mismatch where dwelling availability and affordability play an important role? (NOTE: LMD has have questions about the results) Are there social housing programs for lower income households?

Even if we are on the right direction to develop human-centred urban areas by facilitating caring activities, the chrono-urbanism needs to set the ground on aspects that specifically target caring activities, given the multiple impacts it has on everyday mobility, the unequal distribution of tasks among the population and its related socioeconomic consequences. Therefore, it is necessary to provide alternatives to measure progress on this important issue. City planning should respond to these issues to promote equity. On the other hand, an intersectional approach to understand how different personal identities interact in our daily day activities in the cities is necessary.
